Delay Comparison of Different Switch Architectures

### 2013-11-19 MSReport/MSThesis.tex
first draft

\documentclass{IEEEtran}%[conference]{IEEEtran}%\documentclass[10pt]{article}%{article}{report}{letter}{book}{proc}{slides}%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{skull}

@@

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{Stephan Adams and Libin Jiang}
%\IEEEauthorblockA{Department of Electrical Engineering and Computer Science\\
%University of California, Berkeley\\
%\{shadams, jnchang\}@eecs.berkeley.edu}
\author{\IEEEauthorblockN{S. H. Adams, L. Huang, A. Parekh, and J. Walrand \\}}
%\IEEEauthorblockA{Department of Electrical Engineering and Computer Science\\
%University of California, Berkeley\\
%\{shadams,jnchang\}@eecs.berkeley.edu}}

@@

{\abstract
 We compare the delays experienced by packet flows when transmitted using different scheduling algorithms across a crossbar switch.  The two scheduling algorithms we consider are iterative Slip \cite{McKeown} and QCSMA \cite{Libin}.  We first compare them under the assumptions that all packets have the same length, and then under the assumption that the half the packets have the maximum allowable length and half have the minimum allowable length.  Our findings suggest that the variation has a non-negligible effect on throughput vs delay results, but doesn't have a large qualitative effect on TCP performance.  While QCSMA shows an improved performance over SLIP in open loop simulations, the effect does not translate to better closed loop performance.
}

\section{Introduction}
Computer networks have made their way to the core of our modern lifestyles.  Almost any information can (and often must) be found on the internet.  Many of the services people use on a daily basis rely on distributed processing of requests housed in enormous datacenters scattered across the country and the world.  All these networks require good routing technologies to be possible.

What is the best way to measure the efficacy of a router?  Two obvious metrics are the throughput and delay characteristics of the switches.  Traditionally routers have been evaluated based on their throughput performance.  This is an excellent metric as in the internet routing paths are subject to too much uncertainty to hope that optimizing the delays in a single switch to have a large influence on the roundtrip time of your packets.

On the other hand, in a datacenter setting the paths are no longer uncertain, as the whole network is controlled by a single entity, so delay again becomes a metric that is worth considering.  The other reason that delay optimization ) is a good idea is that the distributed nature of the applications run on datacenters is often very delay sensitive.  User facing applications often ned to return results before a given time threshold or the work must be discarded.%For instance in a google data center any search query must be returned within 300 milliseconds? (CITE SOMEONE) or the work done must be discarded.

Although delay is clearly a very important metric to consider in these settings, it is not as readily analyzed as throughput, which makes it more difficult to give a reliable a design recommendation.  In order to model the effects of different switch schedulers we have built a network simulator using assumptions meant to mimic a datacenter environment.  We have compared the performance of two relevant scheduling architectures in the hopes of demonstrating that provable throughput optimality can lead to an improvement in delay performance when switches are subjected to high loads.

\section{Switch Schedulers}
We consider two different scheduling algorithms used to route packets across a crossbar based switch.  Iterative Slip -- a practical scheme design by Nick McKeown to solve the issues of head of line delay \cite{McKeown}, and QCSMA -- a probabilistic scheduling policy developed and optimized to be full throughput by Libin Jiang \cite{Libin}.

Our assumptions on the hardware for both our switches is the same, the only differentiating component is the scheduler which determines which set of packets will be transmitted at each transmission opportunity.  A good scheduler can support an arbitrarily high load, and provides a low delay per packet.  In practice the scheduler will incur some overhead because it takes time to generate a feasible transmission schedule.  The schedulers use an (ideally) small number of contention slots at the beginning of each packet transmission opportunity during which the inputs and the outputs can exchange information to determine the best schedule.  The communication mechanism between the inputs and outputs can be quite powerful in practice.  We assume that each input is connected to each output by a bidirectional one bit line. This allows each output and input to exchange 1 bit in each contention slot.  For the Slip based scheduler the packets are broken into smaller cells so that the schedules can be generated for a given set of cells, after which they are transmitted over the crossbar, and then the process is repeated.  The QCSMA based algorithms do not synchronize scheduling times, rather it is constantly generating new schedules across the unused portion of the crossbar.

@@

SLIP is a very intuitive and practical scheme invented by Nick Mckeown \cite{McKeown}.  It has been implemented in real systems and yields extremely good performance.  It is difficult to prove anything about this scheduler, but simulations show that it performs very well in many key loading regimes.\\

@@

After all the last contention slot the scheduled input and outputs transmit a single packet and update their state, after which the cycle continues.  Note that since the time it takes to transmit a packet depends on its length, there will be a lot of idle time in the system if packet length varies significantly.  In practice this is addressed by breaking packets up into small equal sized cells so scheduling can be done on a cell by cell basis as opposed to a packet by packet basis.  In our simulations we consider both the case of all packets being the same size, and the case in which there is a bimodal distribution of packet sizes.  The overhead necessary to split packets into smaller cells involves adding a header to each cell, and using a larger number of contention slots since they will be needed to schedule per cell size, not packet size.  In our simulations we assumed both of these contributed only a negligible amount to the total delay.

@@

Still ideal QCSMA, is guaranteed to be full throughput.  Unfortunately the delay performance of priority flows will depend on the backlog of the queues which they enter, which penalizes low rate flows.  The ideal QCSMA takes place in continuous time and so is not constrained by the number of contention slots allotted to the scheduler, and thus represents a performance bound for the other schedulers.\\

@@

Note the probability of a collision (two or more inputs requesting access to the same output simultaneously) is 0 when the timer values $X_{s,d}$ are continuous.  In simulations we assume that the timer resolution period takes no time, so this scheduling policy can act as an idealized performance benchmark.  This is impossible to implement, so it is necessary to develop a time slotted version that works in a finite number of discrete of contention slots.  Because we assume uniform packet sizes, and synchronized scheduling, all inputs compete for all outputs at the beginning of each transmission opportunity.  This may negatively impact observed performance.% although this is unimplementable.  Also note that QCSMA is inspired by Carrier Sense Multiple Access protocols which do not take advantage of the more fine grained feedback available in the switch context, and leveraged by SLIP.

@@

In a real system it is not possible to implement continuous time counters, so it is necessary to approximate the continuous time scheduler with a discrete time system.  The scheduling requests for a given output from a given source are made whenever both that output and source are not transmitting any packets and so is not limited to a fixed number of contention slots like that of Slip.  (We can think of it as in every time slot, performing QCSMA on the portion of the crossbar that is idle.)  Slotted QCSMA approximated exponential timers by geometric random variables determining in which time slot the transmission requests should be made.

@@

To compare the delay performance of the different schedulers we built an event driven simulators in C++.  All the graphs were plotted for $32 \times 32$ switches (i.e. 32 inputs and 32 outputs).  Simulation time is discrete with the minimum time step assumed to be the length of one of the bit transmissions of one of the schedulers.   The packet lengths are given in lengths of these time steps.  Because modern crossbars are transmit packets in parallel we assume that it takes 5 clock ticks to transmit a 64 byte packet, and 120 clock ticks to transmit a 1500 byte packet.  For all the Slip simulations we assumed there were $6$ contention slots used to generate the crossbar schedules, but that these calculations were pipelined in such a way as to make the net scheduling was negligible compared to the packet transmission time.  Simulation inputs were flow patterns passing through the switch, which were identical for all switches. %The switch operation was simulated for $10^6$ time steps, in order to reach the steady state switch behavior%during a given experiment in order to avoid  make the comparison between different scheduling policies and module types without worrying about inconsistencies in flow statistics.

@@

In figure \ref{one_size}, all the packets were generated to have one size.  In this scenario Slip performs very well, on par with ideal QCSMA.  Slotted QCSMA incurs the largest delays.

In order to test whether this poor performance was due to the higher collision rate caused by all the packet transmissions beginning and ending almost simultaneously, we performed a second experiment in which we had a bimodal packet distribution over packet lengths.  In this alternate scenario (depicted in figure \ref{variable_size}) we generated packets with an iid probability of being the maximum size (1500 bytes) with probability $1/2$ and the minimum size (64 bytes) with probability $1/2$.  The rationale behind this distribution was to observe that data packets in TCP flows are generally transferring chunks of some larger file (so may as well send the maximum bytes).  For every data packet which makes it through the system, there is a control message, such as an ACK that do not need to be very large so can be modeled as being the minimum data packets.  As shown in figure \ref{variable_size} all the schedulers perform comparably until we reach higher loads.  Above loads of $60\%$ the switch capacity, Slip starts to perform considerably worse than slotted QCSMA.  Interestingly the performance of the slotted QCSMA follows the performance of the ideal version more closely than in the single packet length scenario.

\begin{figure}%[h]
	 \includegraphics[width=90mm]{us_load.eps}
	\caption{Delay vs throughput graphs for loads in a scenario in which all packet have the same length.  Slip performs comparably to the ideal QCSMA scheduler, and outperforms the realistic implementation of QCSMA.  The parameters for the QCSMA based schedulers are taken from table \ref{qcsma_parameters}.} 	
	\label{one_size}
\end{figure}

\begin{figure}%[h]
	 \includegraphics[width=90mm]{vs_load.eps}
	 \caption{Delay vs throughput graphs for loads in a scenario in which every packet has a probability of $1/2$ of being the minimum size and $1/2$ of being the maximum size. With the variable sized packets Slip no longer outperforms the practical version of QCSMA.  QCSMA still performs similarly to the single length scenario.  (Note that the x axis in this graph is different from that in figure \ref{one_size}.) The parameters for the QCSMA based schedulers are taken from table \ref{qcsma_parameters}.} 	
	\label{variable_size}
\end{figure}

@@

\begin{table}[ht] \caption{TCP Parameters} 
\centering 
\begin{tabular}{c c c}
 \hline\hline 
 Switch & $T_{\text{ecn}}$ & $p_{\text{ecn}}$ \\
  [0.5ex] \hline 
 SLIP&10&.5 \\
  QCSMA&12&.5  \\
   ideal  QCSMA&10&.9  \\
   %DCSMA&31&25  \\
  [1ex] \hline 
  \end{tabular}
   \label{ecn_table} 
\end{table}

@@

This does not accurately model a MapReduce job for several reasons.  The biggest departure from a typical MapReduce job is that our type 2 flows do not have similar synchronization or burstiness properties due to the similar completion times of computations in a real MapReduce job.  Furthermore the large delays imposed by the artificial 10 hop distance for the acks is responsible for another difference in our model.  Still we hope that it suffices to give an impression of how a spread and aggregate application might behave.

@@

\begin{figure}%[h]
	 \includegraphics[width=90mm]{all_2_all.eps}
	\caption{This graph shows the delays experienced by a core router with all to all traffic.}
	\label{all_to_all}
\end{figure}

@@

\begin{figure}%[h]
	 \includegraphics[width=90mm]{vs_spread.eps}
	\caption{This graph shows the delay spread of type 1 flows under a simulated spread and aggregate pattern.  Note the difference in average performance (as represented by the red dot).}
	\label{vs_spreadagg}
\end{figure}

@@

\begin{figure}%[h]
	 \includegraphics[width=90mm]{typ_vs.eps}
	\caption{This graph shows the delays experienced by a typical set of datacenter flows.}
	\label{typ_vs}
\end{figure}

\begin{figure}%[h]
	 \includegraphics[width=90mm]{typ_spread_vs.eps}
	\caption{This graph shows the delays experienced by a typical set of datacenter flows with variable packet sizes and two spread aggregate jobs.}
	\label{typ_spread_vs}
\end{figure}

@@

%% \begin{itemize}
%%\item Metric is Delay (makes sense in DataCenter Environments)
%%\item Delay VS Throughput result changes dramatically depending on packet size assumptions
%%\item TCP acts as expected
%%\end{itemize}
%%
%%\begin{itemize}
%%\item Routers are a core piece of our networks
%%\item Need simple scheduling policies, since volume of traffic strongly penalizes per packet overheads 
%%\item In the internet throughput is right metric -- can't make up for other delays in network
%%\item In Data centers have enough control over other switches that delay becomes a good metric.
%%\item Evolution of Routers -- SLIP killed head of Line delays but can't really be analyzed
%%\item QCSMA -- solves NP hard problem, but may cause large delays?
%%\item Question:  Should the optimality of QCSMA not show up in some performance gains over slip?  (In particular wrt delay?)
%%\end{itemize}
%%
%%\section{Schedulers}
%%\begin{itemize}
%%\item How they work
%%\item Design Philosophy
%%\item Expected Properties
%%\end{itemize}
%%
%%\section{Experiments}
%%\begin{itemize}
%%\item Uniform Packet Sizes VS More Realistic Setting
%%\item Parameter Choices
%%\item Load Graphs
%%\item TCP Graphs Choice of Load Patterns
%%\end{itemize}
%%
%%\section{Conclusions}
%%\begin{itemize}
%%\item Differences in Packet Sizes Cannot be Neglected
%%\item QCSMA may perform better than slip when looking at delay optimization
%%\item TCP in datacenters...
%%\end{itemize}
%%
%%\section{About Simulator:}
%%\begin{itemize}
%%\item Event Driven
%%\item variable collections
%%\item have state, events, and update functions
%%\item use memoryless thing
%%\item main loop
%%\item to change things need to change event merge, possibly the state, and which functions are called
%%\item event pictures
%%\end{itemize}
%%
%%\section{Next Step:}
%%\begin{itemize}
%%\item pick suitable parameters
%%\item Build Experiments?
%%\end{itemize}
%%\section{To Do}
%%\begin{itemize}
%%\item Build Simulator: Now -- Oct 15th?
%%\item Perform Experiments: Oct 15th -- November 15th?
%%\item Write Paper: Now -- December 15th?
%%\item Time Line
%%\end{itemize}
%%
%%\section{Simulator Pieces:}
%%\subsection{Crossbar}
%%\subsection{Scheduling Modules}
%%\subsection{Input Interface}
%%\subsection{Packet Generation}
%%\subsubsection{TCP Layer}
%%\subsubsection{Application Layer}
%%\subsection{Experiment Functions}
%%\subsection{Analysis Tools}
%%\subsubsection{Logs}
%%\subsubsection{Data Collection}
%%\subsubsection{Unit Tests}
%%\begin{itemize}
%%\item state is properly initialized
%%\item event merging
%%\item only real schedules are generated
%%\item right averages are attained
%%\item TCP waveforms are right when the channel is predictable
%%\item Event merge.
%%\end{itemize}
%%\subsubsection{Matlab Data Display Tools}
%%\subsection{Setting Parameters}
%%
%%
%%\section{Perform Experiments}
%%
%%Try to mirror the experiments with the previous simulator as much as possible.
%%\subsection{What kind of packet distributions?}
%%Try bimodal distribution with high probability around a big size and small size with slight variations around the means.  Also single size with some slight variations.  Finally try without any of the small variations.
%%\subsection{What kind of Flow patterns?}
%%Start with those we have decided model the datacenter.
%%\subsection{Throughput Vs Delay Graphs}
%%\subsubsection{What are good parameters for QCSMA?}
%%\subsubsection{Number of Slots for SLIP}
%%\subsection{TCP performance}
%%\subsection{Performance Variation as packet sizes become more predictable}
%%
%%\section{Write Paper}
%%\subsection{Would like to show: QCSMA is better than SLIP by x\%}
%%\subsection{Would like to show: Equal Length vs Variable Length Packets makes a big difference!}
%%\subsection{Simulator Documentation}
%%\subsubsection{Events and State}
%%\subsubsection{How to change things}
%%

\label{bib}
\bibliography{MSThesis}{}
\bibliographystyle{plain} 
 \newpage
 %\appendix
  \appendices 
 \section{Simulation Design} \label{code_design}

@@

So to add a component of a simulator it would be necessary to do the following things:
\begin{itemize}
\item Create additional state variable\\
--- This is to facilitate any bookeeping necessary for events generated by this new component
\item Update the Event class to include updates to the new state.\\
--- This consists of two important parts: a modification of variables in the Event class, and an update to the {\it merge()} method, so two events modifying your new state can successfully be merged into one.
\item The creation of a new function to be placed into the next$\_$event() function.\\
--- This new function would have the form of a next$\_${\it new$\_$component}$\_$event() and would calculate the next event that would occur given the current state.
\item Ideally there would also be a control variable allowing for an easy way to remove the component if it is not needed in a simulation.
\end{itemize}
%%
%%\begin{itemize}
%%\item To illustrate what the functions called by next$\_$event() look like we have the example of:
%%\item To add a component to the simulator, need to modify:
%%\item switch statements 
%%\end{itemize}
\subsection{Auxiliary structures}
In addition to the simulation mechanisms there are a few code structures meant to improve the ease of use.  Data is collected by entering it into an object of the Data$\_$Collector class which provides a convenient dump$\_$to$\_$file() method, which dumps the data into a csv file.  Once all the data is recorded in a file, you can easily graph the data using a Matlab script.  
%%\begin{itemize}
%%\item variable collections and switch statements
%%\item Data Collector Class
%%\item Matlab
%%\end{itemize}

>>>>

\documentclass[11pt]{article}%{IEEEtran}%[conference]{IEEEtran}%\documentclass[10pt]{article}%{article}{report}{letter}{book}{proc}{slides}%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{wasysym}
%\usepackage{skull}

@@

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{Stephan Adams and Libin Jiang}
%\IEEEauthorblockA{Department of Electrical Engineering and Computer Science\\
%University of California, Berkeley\\
%\{shadams, jnchang\}@eecs.berkeley.edu}
\author{S. H. Adams, L. Huang, A. Parekh, and J. Walrand }%\author{\IEEEauthorblockN{S. H. Adams, L. Huang, A. Parekh, and J. Walrand \\}}
%\IEEEauthorblockA{Department of Electrical Engineering and Computer Science\\
%University of California, Berkeley\\
%\{shadams,jnchang\}@eecs.berkeley.edu}}

@@

{\abstract
 We compare the delays experienced by packet flows when transmitted using different scheduling algorithms across a crossbar switch.  The two scheduling algorithms we consider are iterative SLIP \cite{McKeown} and QCSMA \cite{Libin}.  We first compare them under the assumptions that all packets have the same length, and then under the assumption that the half the packets have the maximum allowable length and half have the minimum allowable length.  Our findings suggest that the variation in packet length has a non-negligible effect on throughput vs delay results.  Under varying packet lengths and under TCP flow control, QCSMA derived schedulers seem to do marginally better than SLIP.}

\section{Introduction}
Computer networks have made their way to the core of our modern lifestyles.  Almost any information can (and often must) be found on the internet.  Many of the services people use on a daily basis rely on distributed processing of requests housed in enormous datacenters scattered across the country and the world.  All these networks require good routing technologies to be used effectively.

What is the best way to measure the efficacy of a router?  Two obvious metrics are the throughput and delay characteristics of the switches.  Traditionally routers have been evaluated based on their throughput performance.  This is an excellent metric as, in the internet routing, paths are subject to too much uncertainty to hope that optimizing the delays in a single switch to have a large influence on the roundtrip time of your packets.

On the other hand, in a datacenter setting, paths are no longer uncertain, as the whole network is controlled by a single entity.  This allows delay to become a metric worth considering.  The other reason that delay optimization is a metric worth pursuing is because the distributed applications run on datacenters are often very delay sensitive.  User facing applications often ned to return results before a given time threshold or the work must be discarded.%For instance in a google data center any search query must be returned within 300 milliseconds? (CITE SOMEONE) or the work done must be discarded.

Although delay is clearly a very important metric to consider in these settings, it is not as readily analyzed as throughput, which makes it more difficult to give a reliable a design recommendation.  In order to model the effects of different switch schedulers we have built a network simulator using assumptions meant to mimic a datacenter environment.  We have compared the performance of two relevant scheduling architectures in the hopes of demonstrating that schedulers that are provably throughput optimal have better delay performance when switches are subjected to high loads.

\section{Switch Schedulers}
We consider two different scheduling algorithms used to route packets across a crossbar based switch.  Iterative SLIP -- a practical scheme design by Nick McKeown to solve the issues of head of line delay \cite{McKeown}, and QCSMA -- a probabilistic scheduling policy developed and optimized to be full throughput by Libin Jiang \cite{Libin}.

Our assumptions on the hardware for both our switches is the same, the only differentiating component is the scheduler which determines which set of packets will be transmitted at each transmission opportunity.  A good scheduler can support an arbitrarily high load, and provides a low delay per packet.  In practice the scheduler will incur some overhead because it takes time to generate a feasible transmission schedule.  The schedulers use an (ideally) small number of contention slots at the beginning of each packet transmission opportunity during which the inputs and the outputs can exchange information to determine the best schedule.  The communication mechanism between the inputs and outputs can be quite powerful in practice.  We assume that each input is connected to each output by a bidirectional one bit line. This allows each output and input to exchange 1 bit in each contention slot.  For the SLIP based scheduler the packets are broken into smaller cells so that the schedules can be generated for a given set of cells, after which they are transmitted over the crossbar, and then the process is repeated.  The QCSMA based algorithms do not synchronize scheduling times, rather it is constantly generating new schedules across the unused portion of the crossbar.

@@

Iterative SLIP is a very intuitive and practical scheme invented by Nick Mckeown \cite{McKeown}.  It has been implemented in real systems and yields extremely good performance.  It is difficult to prove anything about this scheduler, but simulations show that it performs very well in many key loading regimes.\\

@@

After the last contention slot, the scheduled input and outputs transmit a single packet and update their state.  Once the packets have been successfully transmitted, the cycle continues.  Note that since the time it takes to transmit a packet depends on its length, there will be a lot of idle time in the system if packet length varies significantly.  In practice this is addressed by breaking packets up into small equal sized cells so scheduling can be done on a cell by cell basis as opposed to a packet by packet basis.  In our simulations we consider both the case of all packets being the same size, and the case in which there is a bimodal distribution of packet sizes.  The overhead necessary to split packets into smaller cells involves adding a header to each cell, and using a larger number of contention slots since they will be needed to schedule per cell size, not packet size.  In our simulations we assumed both of these contributed only a negligible amount to the total delay.

@@

Still ideal QCSMA, is guaranteed to be full throughput.  Unfortunately the delay performance of priority flows will depend on the backlog of the queues which they enter, which penalizes low rate flows.  The ideal QCSMA takes place in continuous time: so it can run arbitrarily fast, as well as being immune to collisions, and thus represents a performance bound for the other schedulers.\\

@@

Note the probability of a collision (two or more inputs requesting access to the same output simultaneously) is 0 when the timer values $X_{s,d}$ are continuous.  In simulations we assume that the timer resolution period takes no time, so this scheduling policy can act as an idealized performance benchmark.  This is impossible to implement, so it is necessary to develop a time slotted version that works in a finite number of discrete of contention slots.% although this is unimplementable.  Also note that QCSMA is inspired by Carrier Sense Multiple Access protocols which do not take advantage of the more fine grained feedback available in the switch context, and leveraged by SLIP.

@@

In a real system it is not possible to implement continuous time counters, so it is necessary to approximate the continuous time scheduler with a discrete time system.  The scheduling requests for a given output from a given source are made whenever both that output and source are not transmitting any packets and so is not limited to a fixed number of contention slots like that of SLIP.  (We can think of it as in every time slot, performing QCSMA on the portion of the crossbar that is idle.)  Slotted QCSMA approximated exponential timers by geometric random variables determining in which time slot the transmission requests should be made.

@@

To compare the delay performance of the different schedulers we built an event driven simulators in C++.  All the graphs were plotted for $32 \times 32$ switches (i.e. 32 inputs and 32 outputs).  Simulation time is discrete with the minimum time step assumed to be the length of one of the bit transmissions of one of the schedulers.   The packet lengths are given in lengths of these time steps.  Because modern crossbars are transmit packets in parallel we assume that it takes 5 clock ticks to transmit a 64 byte packet, and 120 clock ticks to transmit a 1500 byte packet.  For all the SLIP simulations we assumed there were $6$ contention slots used to generate the crossbar schedules, but that these calculations were pipelined in such a way as to make the net scheduling was negligible compared to the packet transmission time.  Simulation inputs were flow patterns passing through the switch, which were identical for all switches. %The switch operation was simulated for $10^6$ time steps, in order to reach the steady state switch behavior%during a given experiment in order to avoid  make the comparison between different scheduling policies and module types without worrying about inconsistencies in flow statistics.

@@

In figure \ref{one_size}, all the packets were generated to have one size.  In this scenario SLIP performs very well, on par with ideal QCSMA.  Slotted QCSMA incurs the largest delays.

In order to test whether this poor performance was due to the higher collision rate caused by all the packet transmissions beginning and ending almost simultaneously, we performed a second experiment in which we had a bimodal packet distribution over packet lengths.  In this alternate scenario (depicted in figure \ref{variable_size}) we generated packets with an iid probability of being the maximum size (1500 bytes) with probability $1/2$ and the minimum size (64 bytes) with probability $1/2$.  The rationale behind this distribution was to observe that data packets in TCP flows are generally transferring chunks of some larger file (so may as well send the maximum bytes).  For every data packet which makes it through the system, there is a control message, such as an ACK that do not need to be very large so can be modeled as being the minimum data packets.  As shown in figure \ref{variable_size} all the schedulers perform comparably until we reach higher loads.  Above loads of $60\%$ the switch capacity, SLIP starts to perform considerably worse than slotted QCSMA.  Interestingly the performance of the slotted QCSMA follows the performance of the ideal version more closely than in the single packet length scenario.

\begin{figure}%[h]
	 \includegraphics[width=90mm]{us_load.eps}
	\caption{Delay vs throughput graphs for loads in a scenario in which all packet have the same length.  SLIP performs comparably to the ideal QCSMA scheduler, and outperforms the realistic implementation of QCSMA.  The parameters for the QCSMA based schedulers are taken from table \ref{qcsma_parameters}.} 	
	\label{one_size}
\end{figure}

\begin{figure}%[h]
	 \includegraphics[width=90mm]{vs_load.eps}
	 \caption{Delay vs throughput graphs for loads in a scenario in which every packet has a probability of $1/2$ of being the minimum size and $1/2$ of being the maximum size. With the variable sized packets SLIP no longer outperforms the practical version of QCSMA.  QCSMA still performs similarly to the single length scenario.  (Note that the x axis in this graph is different from that in figure \ref{one_size}.) The parameters for the QCSMA based schedulers are taken from table \ref{qcsma_parameters}.} 	
	\label{variable_size}
\end{figure}

@@

\begin{table}[ht] \caption{TCP Parameters} 
\centering
\begin{tabular}{c c c}
 \hline\hline 
 Switch & $T_{\text{ecn}}$ & $p_{\text{ecn}}$ \\
  [0.5ex] \hline 
 SLIP&10&.5 \\
  QCSMA&12&.5  \\
   ideal  QCSMA&10&.9  \\
  [1.0ex] \hline 
  \end{tabular}
   \label{ecn_table} 
\end{table}

@@

This does not accurately model a MapReduce job for several reasons.  The biggest departure from a typical MapReduce job is that our type 1 flows do not have similar synchronization or burstiness properties due to the similar completion times of computations in a real MapReduce job.  Furthermore the large delays imposed by the artificial 10 hop distance for the acks is responsible for another difference in our model.  Still we hope that it suffices to give an impression of how a spread and aggregate application might behave.

@@

\begin{figure}%[h]
	 \includegraphics[width=\textwidth]{all_2_all.eps}
	\caption{This graph shows the delays experienced by a core router with all to all traffic.}
	\label{all_to_all}
\end{figure}

@@

\begin{figure}%[h]
	 \includegraphics[width=\textwidth]{vs_spread.eps}
	\caption{This graph shows the delay spread of type 1 flows under a simulated spread and aggregate pattern.  Note the difference in average performance (as represented by the red dot).}
	\label{vs_spreadagg}
\end{figure}

@@

\begin{figure}%[h]
	 \includegraphics[width=\textwidth]{typ_vs.eps}
	\caption{This graph shows the delays experienced by a typical set of datacenter flows.}
	\label{typ_vs}
\end{figure}

\begin{figure}%[h]
	 \includegraphics[width=\textwidth]{typ_spread_vs.eps}
	\caption{This graph shows the delays experienced by a typical set of datacenter flows with variable packet sizes and two spread aggregate jobs.}
	\label{typ_spread_vs}
\end{figure}

@@

%% \begin{itemize}
%%\item Metric is Delay (makes sense in DataCenter Environments)
%%\item Delay VS Throughput result changes dramatically depending on packet size assumptions
%%\item TCP acts as expected
%%\end{itemize}
%%
%%\begin{itemize}
%%\item Routers are a core piece of our networks
%%\item Need simple scheduling policies, since volume of traffic strongly penalizes per packet overheads 
%%\item In the internet throughput is right metric -- can't make up for other delays in network
%%\item In Data centers have enough control over other switches that delay becomes a good metric.
%%\item Evolution of Routers -- SLIP killed head of Line delays but can't really be analyzed
%%\item QCSMA -- solves NP hard problem, but may cause large delays?
%%\item Question:  Should the optimality of QCSMA not show up in some performance gains over SLIP?  (In particular wrt delay?)
%%\end{itemize}
%%
%%\section{Schedulers}
%%\begin{itemize}
%%\item How they work
%%\item Design Philosophy
%%\item Expected Properties
%%\end{itemize}
%%
%%\section{Experiments}
%%\begin{itemize}
%%\item Uniform Packet Sizes VS More Realistic Setting
%%\item Parameter Choices
%%\item Load Graphs
%%\item TCP Graphs Choice of Load Patterns
%%\end{itemize}
%%
%%\section{Conclusions}
%%\begin{itemize}
%%\item Differences in Packet Sizes Cannot be Neglected
%%\item QCSMA may perform better than SLIP when looking at delay optimization
%%\item TCP in datacenters...
%%\end{itemize}
%%
%%\section{About Simulator:}
%%\begin{itemize}
%%\item Event Driven
%%\item variable collections
%%\item have state, events, and update functions
%%\item use memoryless thing
%%\item main loop
%%\item to change things need to change event merge, possibly the state, and which functions are called
%%\item event pictures
%%\end{itemize}
%%
%%\section{Next Step:}
%%\begin{itemize}
%%\item pick suitable parameters
%%\item Build Experiments?
%%\end{itemize}
%%\section{To Do}
%%\begin{itemize}
%%\item Build Simulator: Now -- Oct 15th?
%%\item Perform Experiments: Oct 15th -- November 15th?
%%\item Write Paper: Now -- December 15th?
%%\item Time Line
%%\end{itemize}
%%
%%\section{Simulator Pieces:}
%%\subsection{Crossbar}
%%\subsection{Scheduling Modules}
%%\subsection{Input Interface}
%%\subsection{Packet Generation}
%%\subsubsection{TCP Layer}
%%\subsubsection{Application Layer}
%%\subsection{Experiment Functions}
%%\subsection{Analysis Tools}
%%\subsubsection{Logs}
%%\subsubsection{Data Collection}
%%\subsubsection{Unit Tests}
%%\begin{itemize}
%%\item state is properly initialized
%%\item event merging
%%\item only real schedules are generated
%%\item right averages are attained
%%\item TCP waveforms are right when the channel is predictable
%%\item Event merge.
%%\end{itemize}
%%\subsubsection{Matlab Data Display Tools}
%%\subsection{Setting Parameters}
%%
%%
%%\section{Perform Experiments}
%%
%%Try to mirror the experiments with the previous simulator as much as possible.
%%\subsection{What kind of packet distributions?}
%%Try bimodal distribution with high probability around a big size and small size with slight variations around the means.  Also single size with some slight variations.  Finally try without any of the small variations.
%%\subsection{What kind of Flow patterns?}
%%Start with those we have decided model the datacenter.
%%\subsection{Throughput Vs Delay Graphs}
%%\subsubsection{What are good parameters for QCSMA?}
%%\subsubsection{Number of Slots for SLIP}
%%\subsection{TCP performance}
%%\subsection{Performance Variation as packet sizes become more predictable}
%%
%%\section{Write Paper}
%%\subsection{Would like to show: QCSMA is better than SLIP by x\%}
%%\subsection{Would like to show: Equal Length vs Variable Length Packets makes a big difference!}
%%\subsection{Simulator Documentation}
%%\subsubsection{Events and State}
%%\subsubsection{How to change things}
%%

\label{bib}
\bibliography{MSThesis}{}
\bibliographystyle{plain} 
 \newpage
\appendix
%  \appendices 
 \section{Simulation Design} \label{code_design}

@@

So to add a component of a simulator it would be necessary to do the following things:
\begin{itemize}
\item Create additional state variable\\
--- This is to facilitate any bookeeping necessary for events generated by this new component
\item Update the Event class to include updates to the new state.\\
--- This consists of two important parts: a modification of variables in the Event class, and an update to the {\it merge()} method, so two events modifying your new state can successfully be merged into one.
\item The creation of a new function to be placed into the next$\_$event() function.\\
--- This new function would have the form of a next$\_${\it new$\_$component}$\_$event() and would calculate the next event that would occur given the current state.
\item Ideally there would also be a control variable allowing for an easy way to remove the component if it is not needed in a simulation.
\end{itemize}
%%
%%\begin{itemize}
%%\item To illustrate what the functions called by next$\_$event() look like we have the example of:
%%\item To add a component to the simulator, need to modify:
%%\item switch statements 
%%\end{itemize}
\subsection{Auxiliary structures}
In addition to the simulation mechanisms there are a few code structures meant to improve the ease of use.  Data is collected by entering it into an object of the Data$\_$Collector class which provides a convenient dump$\_$to$\_$file() method, which dumps the data into a csv file.  Once all the data is recorded in a file, you can easily graph the data using a Matlab script.\\

Because of the size of the simulator, there are many sets of variables floating around which may make it difficult to follow what is relevant to what.  In order to combat this, the simulator has collected the different variables parametrizing the different components in structs such as the TCP$\_$State or the Scheduler$\_$Parameters struct, defined at the beginning of the code so that it is easy to follow what values need to be defined in order to make the simulator function.\\

Another issue with understanding the simulator code is that it contains components that can be added or removed (such as the TCP layer or the current choice of scheduler).  What simulator components are currently active is parametrized by the Simulation$\_$Parameters struct, and at the relevant places in the code, there is a switch statement or a an if statement which switches between the relevant components.  Searching for the variables collected in Simulation\_Parameters should quickly reveal any places that have been modified or need to modified by the particular component.
%%\begin{itemize}
%%\item variable collections and switch statements
%%\item Data Collector Class
%%\item Matlab
%%\end{itemize}
